My title

### 2024-03-11 paper/paper.qmd
extra data removed, and new data of charts added

library(tidyverse)
library(palmerpenguins)
library(readxl)
library(here)

deaths_data <- read.csv(here("data/analysis_data/cleaned_file.csv"))
air_quality_data <- read.csv(here("data/analysis_data/cleaned_air_data.csv"))
```

@@

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

@@

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

# analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
# 
# analysis_data |> 
#   ggplot(aes(x = width, y = length)) +
#   geom_point(alpha = 0.8) +
#   theme_minimal() +
#   labs(x = "Wing width (mm)",
#        y = "Wing length (mm)")
```

@@

first_model <-
  readRDS(file = here::here("models/first_model.rds"))

poisson_model <-
  readRDS(file = here::here("models/poisson_model.rds"))

causes_model <-
  readRDS(file = here::here("models/causes_model.rds"))
```

@@

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  #statistic = "mad",
  fmt = 2
)

modelsummary::modelsummary(
  list(
    "First model" = poisson_model
  ),
  #statistic = "mad",
  fmt = 2
)

modelsummary::modelsummary(
  list(
    "First model" = causes_model
  ),
  #statistic = "mad",
  fmt = 2
)
```

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

# Create the posterior predictive check plot
pp_plot <- pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

# Create the posterior vs prior plot
posterior_plot <- posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

>>>>

library(tidyverse)
library(palmerpenguins)
library(here)
library(ggplot2)

#Reading the data
deaths_data <- read.csv(here("data/analysis_data/cleaned_file.csv"))
air_quality_data <- read.csv(here("data/analysis_data/cleaned_air_data.csv"))
avg_PM_conc_data <- read.csv(here("data/raw_data/avg PM2.5 conc data.csv"))
peak_PM_conc_data <- read.csv(here("data/raw_data/peak PM2.5 conc data.csv"))
avg_PM_conc_data <- avg_PM_conc_data %>% rename(Year = Category)
peak_PM_conc_data <- peak_PM_conc_data %>% rename(Year = Category)

# Assuming 'cleaned_air_data' is your air quality dataset and 'mortality_data' is your dataset of deaths.

library(tidyverse)
library(ggplot2)

# Ensure you have the correct column names by checking the names of your data frames
colnames(air_quality_data)
colnames(deaths_data)

# Merge the datasets
combined_data <- left_join(deaths_data, air_quality_data, by = "year")

head(combined_data)

# 2. Mortality Trend Plot for a single cause
# Repeat this plot changing the 'Cause' filter for each cause you are interested in.
ggplot(combined_data %>% filter(cause == 'All other forms of chronic ischemic heart disease'), 
       aes(x = year, y = total_deaths)) +  # Note the use of single quotes and backticks
  geom_line() +
  labs(title = "Trend in Deaths Due to Chronic Ischemic Heart Disease",
       x = "Year",
       y = "Total Deaths") +
  theme_minimal()

# 3. Relationship between Air Quality and Mortality Rates
# This could be a scatter plot with a line representing the trend.
ggplot(combined_data, aes(x = original_value, y = total_deaths, color = cause)) +  # Make sure the column name matches
  geom_point() +
  geom_smooth(method = "lm") +  # Add a linear trend line
  facet_wrap(~cause) +  # Create separate plots for each cause
  labs(title = "Relationship between Air Quality and Mortality Rates",
       x = "Air Quality Health Index",
       y = "Total Deaths") +
  theme_minimal()
```

@@

@@

@@

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

@@

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@
